---
title: "Average temperature in Brazilian cities over the last 100 years"
author: "Jaime Pulgar"
date: "5/19/2022"
output: 
  html_document:
    self_contained: false
---

## Dataset:
## Berkeley Earth
## http://berkeleyearth.org/data
## TemperaturasGlobais.csv ~ 78 MB (zip) ~ 530 MB (unzip)

## Installing and loading packages

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(scales)
library(data.table)

```

## Loading the data (Using a timer to compare the loading time with 
## defferent funtions)

## Using the funtion read.csv2()
```{r echo=TRUE}

system.time(df_teste1 <- read.csv2("TemperaturasGlobais.csv"))

```

## Using the funtion read.table()
```{r echo=TRUE}

system.time(df_teste2 <- read.table("TemperaturasGlobais.csv"))

```

## Using the funtion fread()
```{r echo=TRUE}

system.time(df <- fread("TemperaturasGlobais.csv"))

```

## Creating subsets of the loaded data
```{r echo=TRUE}

cidadesBrasil <- subset(df, Country == 'Brazil')
cidadesBrasil <- na.omit(cidadesBrasil)
head(cidadesBrasil)
nrow(df)
nrow(cidadesBrasil)
dim(cidadesBrasil)

```

## Preparation and organization

## Converting the dates
```{r echo=TRUE}
cidadesBrasil$dt <- as.POSIXct(cidadesBrasil$dt,format='%Y-%m-%d')
cidadesBrasil$Month <- month(cidadesBrasil$dt)
cidadesBrasil$Year <- year(cidadesBrasil$dt)
View(cidadesBrasil)

```

## Loading subsets

## Palmas
```{r echo=TRUE}
plm <- subset(cidadesBrasil, City == 'Palmas')
plm <- subset(plm, Year %in% c(1796,1846,1896,1946,1996,2012))

```

## Curitiba
```{r echo=TRUE}
crt <- subset(cidadesBrasil, City == 'Curitiba')
crt <- subset(crt, Year %in% c(1796,1846,1896,1946,1996,2012))

```

## Recife
```{r echo=TRUE}
recf <- subset(cidadesBrasil, City=='Recife')
recf <- subset(recf,Year %in% c(1796,1846,1896,1946,1996,2012))

```

## Building the Plots
```{r echo=TRUE}
p_plm <- ggplot(plm, aes(x = (Month), y = AverageTemperature, color = as.factor(Year))) +
  geom_smooth(se = FALSE,fill = NA, size = 2) +
  theme_light(base_size = 20) +
  xlab("Month")+
  ylab("Average Temperature") +
  scale_color_discrete("") +
  ggtitle("Average Temperature over the years in Palmas") +
  theme(plot.title = element_text(size = 18))

p_crt <- ggplot(crt, aes(x = (Month), y = AverageTemperature, color = as.factor(Year))) +
  geom_smooth(se = FALSE,fill = NA, size = 2) +
  theme_light(base_size = 20) +
  xlab("Month")+
  ylab("Average Temperature") +
  scale_color_discrete("") +
  ggtitle("Average Temperature over the years in Curitiba") +
  theme(plot.title = element_text(size = 18))

p_recf <- ggplot(recf, aes(x = (Month), y = AverageTemperature, color = as.factor(Year))) +
  geom_smooth(se = FALSE,fill = NA, size = 2) +
  theme_light(base_size = 20) +
  xlab("Month")+
  ylab("Average Temperature") +
  scale_color_discrete("") +
  ggtitle("Average Temperature over the years in  Recife") +
  theme(plot.title = element_text(size = 18))

```

# Plotting
```{r cars}
p_plm
p_crt
p_recf

```